Forward Modeling Galaxy Size in a Cosmological Context

### 2017-10-03 paper/galsize_paper.tex
Updated discussion of clustering ratio figure and its caption

In Figure \ref{fig:clustering_ratio_upshot} we present new measurements of the $\rhalf-$dependence of projected galaxy clustering, $\wproj(\rproj).$ Because galaxy clustering has well-known dependence upon $\mstar$ that is not the subject of this work, we wish to remove this influence and focus purely on the relationship between $\rhalf$ and $\wproj(\rproj).$ To do so, we determine the value $\langle\rhalf\vert\mstar\rangle$ by computing a sliding median of $\rhalf,$ calculated using a window of width $N_{\rm gal}=1000.$ Each galaxy is categorized as either ``large" or ``small" according to whether it is above or below the median value appropriate for its stellar mass. For any $\mstar-$threshold sample, the SMF of the ``large" and ``small" subsamples are identical, by construction.

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/penultimate_clustering_ratios_no_orphans.pdf}
\caption{
{\bf $\rhalf-$dependence of galaxy clustering.}
Points with error bars show new SDSS measurements of the $\rhalf-$dependence of projected galaxy clustering, $\wproj,$ compared to predictions by the model tuned to the measurements shown in Fig.~\ref{fig:scatter_plot}. We define a galaxy as ``large" or ``small" according to whether it is above or below the median size for its stellar mass, so that in each panel, the SMF of the ``large" and ``small" subsamples are identical, as described in the text. The y-axis shows {\em clustering strength ratios,} so that, for example, a y-axis value of $-0.5$ corresponds to small galaxies being $50\%$ more strongly clustered than large galaxies of comparable stellar mass. Each panel shows results separately for a different volume-limited $\mstar-$threshold samples. See \S\ref{subsec:model} for a description of each model.}
\label{fig:clustering_ratio_upshot}
\end{figure}
%-----------------------------------------------------------------------------------------------------

The salient feature of the clustering ratio measurements is that they are negative: small galaxies cluster more strongly than large galaxies of the same stellar mass, a new result. This feature also holds true for model galaxies. This result may be surprising, since $\rhalf\propto\rvir,$ halo mass $\rvir\propto\mhalo^{1/3},$ and clustering strength increases with $\mvir.$ Based on this simple argument, one would expect the opposite trend to the measurements shown here.

A straightforward resolution to this puzzle is shown in Figure \ref{fig:censatsizehist}.

@@

%We can understand the origin of the shape of the clustering strength ratios shown in Figure~\ref{fig:clustering_ratio_upshot} through a series of straightforward ``shuffle tests". The points with error bars and solid curves in each panel of Fig.\ref{fig:shuffle} are identical to those appearing in Fig.\ref{fig:clustering_ratio_upshot}, although the dynamic range of the y-axis has changed to accommodate the additional curves.

%We generated each additional curve by variations on the following exercise. In logarithmic bins of $\mpeak$ with width of $0.1$dex, we randomly shuffle the modeled $\rhalf$ values of some particular subpopulation of model galaxies that reside in subhalos in the $\mpeak$ bin. After repeating this shuffle over the entire range of $\mpeak,$ we then re-divide model galaxies into ``large" and ``small" populations, and remeasure the clustering strength ratios. By shuffling sizes amongst different subpopulations, we can gain an understanding of the subhalo properties that contribute to the characteristic shape of the clustering ratios.

%In the {\em brown, dotted} curves labeled ``all-gal scramble", we shuffle the sizes of {\em all} model galaxies in the $\mpeak$ bin, so that in the resulting mock, galaxy size has no correlation with any subhalo property beyond $\mpeak.$ In particular, the ``all-gal scramble" mock erases the possible influence of a correlation between galaxy size and the redshift at which the subhalo attains peak mass. The resulting clustering strength ratios become exactly as predicted by the naive expectation sketched at the conclusion of \S\ref{subsec:predictions}: $\rhalf\propto\rvir\propto\mhalo^{1/3},$ and so large galaxies clustering more strongly relative to small galaxies of comparable stellar mass.

%Statistically, there are actually two distinct effects at work in the ``all-gal scramble".  Central and satellite galaxies of similar $\mpeak$ have their sizes shuffled amongst each other, and the potential influence of central galaxy assembly bias is also erased. We parse these separate effects by exploring three variations on the shuffle tests, described below.

%For the {\em cyan dashed} curve labeled ``all-cen-scramble", we only shuffle sizes amongst present-day host halos, i.e., amongst subhalos for which Rockstar {\tt upid} equals -1. For the {\em black dashed} curve labeled ``true-cen-scramble", we only shuffle amongst host halos that have never passed within the virial radius of a larger halo throughout their entire assembly history, i.e., we only shuffle amongst non-backsplash host halos. Finally, the {\em purple dot-dashed} curve labeled ``sat-scramble", we only shuffle sizes amongst satellite galaxies, i.e., subhalos that are presently located inside the virial radius of a larger halo.

% shown in Figure \ref{fig:shuffle}.

%In light of this explanation, it is natural to ask whether this simple feature {\em alone} is all that is needed for any model of $\rhalf$ to achieve this level of success at predicting galaxy clustering on small and large scales. We address this question in \S~\ref{subsec:tests2} below, finding that the answer is no: the reasonably correct magnitude and scale-dependence of the clustering ratios predicted by our model is a non-trivial result that places tight constraints on the post-infall physics of satellite galaxies.

>>>>

In Figure \ref{fig:clustering_ratio_upshot} we present new measurements of the $\rhalf-$dependence of projected galaxy clustering, $\wproj(\rproj).$ Because galaxy clustering has well-known dependence upon $\mstar$ that is not the subject of this work, we wish to remove this influence and focus purely on the relationship between $\rhalf$ and $\wproj(\rproj).$ To do so, we determine the value $\langle\rhalf\vert\mstar\rangle$ by computing a sliding median of $\rhalf,$ calculated using a window of width $N_{\rm gal}=1000.$ Each galaxy is categorized as either ``large" or ``small" according to whether it is above or below the median value appropriate for its stellar mass. Using this technique, we stress that for any $\mstar-$threshold sample, the SMF of the ``large" and ``small" subsamples are identical, by construction.

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/penultimate_clustering_ratios_no_orphans.pdf}
\caption{
{\bf $\rhalf-$dependence of galaxy clustering.}
Points with error bars show new SDSS measurements of the $\rhalf-$dependence of projected galaxy clustering, $\wproj.$ We define a galaxy as ``large" or ``small" according to whether it is above or below the median size for its stellar mass, so that in each panel, the stellar mass functions of the ``large" and ``small" subsamples are identical, as described in the text. The y-axis shows {\em clustering strength ratios,} so that, for example, a y-axis value of $-0.5$ corresponds to small galaxies being $50\%$ more strongly clustered than large galaxies of comparable stellar mass. Each panel shows results separately for different volume-limited $\mstar-$threshold samples; predictions of three different models are shown in each panel. See \S\ref{subsec:model} for a description of each model.}
\label{fig:clustering_ratio_upshot}
\end{figure}
%-----------------------------------------------------------------------------------------------------

The salient feature of the clustering ratio measurements is that they are negative: small galaxies cluster more strongly than large galaxies of the same stellar mass, a new result. This feature also holds true for galaxies predicted by the $\rvir-$only model. This result may be surprising, since $\rhalf\propto\rvir,$ halo mass $\rvir\propto\mhalo^{1/3},$ and clustering strength increases with $\mvir.$ Based on this simple argument, one would expect the opposite trend to the measurements shown here.

A straightforward resolution to this puzzle is shown in Figure \ref{fig:censatsizehist}, which compares the $\rhalf$ distributions of central, satellite, and splashback galaxies with the same halo mass $\mpeak\approx10^{12}\msun.$ A ``splashback" galaxy is defined as a present-day central that used to be a satellite, i.e., its main progenitor halo passed inside the virial radius of a larger halo at some point in its past history. On the other hand, we define a ``true central" as a galaxy that has never been a satellite.

In the $\rvir-$only model, satellite and splashback galaxies are smaller than centrals of the same halo mass due to the physical size of their halo being smaller at earlier times $\zpeak.$ There are two distinct reasons why this feature results in small galaxies being more strongly clustered relative to larger galaxies of the same mass. First, satellite galaxies statistically occupy higher mass host halos that are more strongly clustered. In models where satellites are smaller than centrals, for any given $\mstar-$threshold the ``small" subsample will naturally have a higher satellite fraction, resulting in a negative clustering ratio as seen in SDSS data. Second, at fixed mass, halos of $L_\ast$ galaxies that form earlier are more strongly clustered, a phenomenon commonly known as {\em halo assembly bias}. Since splashback halos are typically earlier forming than true centrals, then models where splashback halos host smaller-than-average galaxies will naturally predict negative clustering ratios.

@@
